---
title: "Developing Data Products - R Markdown and Leaflet"
author: "Ricardo De Gouveia"
date: "22/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R Markdown and Leaflet


Create a web page using R Markdown that features a map created with Leaflet.

Host your webpage on either GitHub Pages, RPubs, or NeoCities.

Your webpage must contain the date that you created the document, and it must contain a map created with Leaflet. We would love to see you show off your creativity!

## Load Library

```{r}
library(leaflet)
```

# Mapping Chandler/Gilbert High Schools

```{r}
highSchools <- data.frame(name=c("Basha HS", "Perry HS", "Chandler HS", "Hamilton HS"),
                        enrollment = c(2623, 3542, 3241, 4100),
                        lat = c(33.2209, 33.2619, 33.3089, 33.252912),
                        lng = c(-111.7592, -111.7478, -111.8421, -111.842714))

hsMascotIcons <- iconList(
  basha = makeIcon("https://www.cusd80.com/cms/lib/AZ01001175/Centricity/Template/GlobalAssets/images///logos/bhs-logo.png", iconWidth = 31, iconHeight = 31),
  perry = makeIcon("https://www.cusd80.com/cms/lib/AZ01001175/Centricity/Template/GlobalAssets/images///logos/phs-logo.png", iconWidth = 31, iconHeight = 31)               
#  chandler = makeIcon("https://www.cusd80.com/cms/lib/AZ01001175/Centricity/Template/GlobalAsse#ts/images///logos/chs-logo.png", iconWidth = 31, iconHeight = 31)
#  "https://www.cusd80.com/cms/lib/AZ01001175/Centricity/Template/GlobalAsse#ts/images///logos/hhs-logo.png",
)

hsMascotIcons[c("basha", "perry")]
```

```{r}
# lat lng data
df <- sp::SpatialPointsDataFrame(
  cbind(
    (runif(20) - .5) * 10 - 111.5000,  # lng
    (runif(20) - .5) * 3.8 + 33.1000  # lat
  ),
  data.frame(type = factor(
    ifelse(runif(20) > 0.75, "basha", "perry"),
    c("perry", "basha")
  ))
)

hsSites <- c(
  "<a href='https://www.cusd80.com/Domain/457'>Basha HS</a>",
  "<a href='https://www.cusd80.com/Domain/3255'>Perry HS</a>",
  "<a href='https://www.cusd80.com/Domain/913'>Chandler HS</a>",
  "<a href='https://www.cusd80.com/Domain/2039'>Hamilton HS</a>"
)

highSchools %>%
  leaflet() %>%
  addTiles() %>%
  addMarkers(icon = hsMascotIcons, popup = hsSites)
```
