---
title: "Developing Data Products - R Markdown Presentation & Plotly"
author: "Ricardo De Gouveia"
date: "23/8/2020"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## R Markdown Presentation & Plotly

Create a web page presentation using R Markdown that features a plot created with Plotly. Host your webpage on either GitHub Pages, RPubs, or NeoCities. Your webpage must contain the date that you created the document, and it must contain a plot created with Plotly. We would love to see you show off your creativity!

## R code

```{r, echo=TRUE, message=FALSE, warning=FALSE}
library(plotly)
library(ggplot2)
data("mtcars")
head(mtcars)

```

\pagebreak


## Now a Scatter Plot

```{r, message=FALSE,warning=FALSE}
plot_ly(mtcars, x = ~hp, y = ~mpg, color = ~cyl, type = "scatter", mode = "markers")
```


\pagebreak

## Now a BoxPlot

```{r, message=FALSE,warning=FALSE}
# create factors with value labels
mtcars$cyl <- factor(mtcars$cyl,levels=c(4,6,8), labels=c("4cyl","6cyl","8cyl"))

plot_ly(mtcars, x = ~mpg, y = ~cyl, color = ~cyl, type = "box", mode = "markers")
```